---
title: "#30DayMapChallenge"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

# Intro

Very quick maps using open and R. For fun and exploration. #30DayMapChallenge

# Day 1: Points

Get points from OpenStreetMap and display

```{r, echo=F, warning=F, message=F}
#  install.packages(c("tidyverse", "osmdata", "sf"))
library(osmdata) 
library(sf)
library(tidyverse)
```

```{r }
#  install.packages(c("tidyverse", "osmdata", "sf"))
library(osmdata) 
library(sf)
library(tidyverse)

glasgow_pubs <-  getbb(place_name = "glasgow uk") %>%
  opq() %>%
  add_osm_feature("amenity", "pub") %>% 
  osmdata_sf()

plot(st_geometry(glasgow_pubs$osm_points$geometry),
     cex = 0.3,
     col = "darkgreen",
     border = "darkgreen",
     main ="Glasgow pubs")

```

## Day 2: Lines


```{r}
#  install.packages(c("tidyverse", "osmdata", "sf"))
library(osmdata) 
library(sf)
library(tidyverse)


glasgow_cyclepath <- getbb(place_name = "glasgow uk") %>%
  opq() %>%
  add_osm_feature("highway", "cycleway") %>% 
  osmdata_sf()


plot(st_geometry(glasgow_cyclepath$osm_lines),
     cex = 0.3,
     col = "red",
     main = "Glasgow cycle paths")


```

# Day 3 Polygons

```{r}
# install.packages(c("devtools", "tidyverse"))
# devtools::install_github("fozy81/scotmaps")
library(scotmaps)
library(tidyverse)

ma <- marine_areas()
plot(st_geometry(ma), col = "lightblue", border = "grey", main = "Marine Areas")



```

# Day 4 Hexagons

```{r}
#  install.packages(c("devtools", "tidyverse"))
library(osmdata) 
library(sf)
library(tidyverse)

bicycles_parking <-  getbb(place_name = "glasgow uk") %>%
  opq() %>%
  add_osm_feature("amenity", "bicycle_parking") %>% 
  osmdata_sf()

glasgow <- getbb(place_name = "glasgow city uk", format_out = "sf_polygon") 

hex <- st_make_grid(glasgow, cellsize = .01, square = FALSE, what = "polygon") 
hex <- st_sf(hex)
hex$id <- seq_along(hex$hex)
hex$glasgow <- 1
hex <- st_join(hex, bicycles_parking$osm_points, left = FALSE) 
hex <-  group_by(hex, id) %>% summarise(sum(glasgow))
hex <- st_sf(hex)

plot(hex["sum(glasgow)"], main = "Glasgow bicycle parking")





```


# Day 5 Blue


```{r}
#  install.packages(c("devtools", "tidyverse"))
library(osmdata) 
library(sf)
library(tidyverse)

blue <-  getbb(place_name = "glasgow uk") %>%
  opq() %>%
  add_osm_feature(key = "name", value = "blue", value_exact = FALSE, match_case = FALSE) %>% 
  osmdata_sf()

blue <- blue$osm_points
blue <- blue[grep("blue", blue$name, ignore.case = T), ]

blue$name <- as.factor(blue$name)
ggplot() +
  geom_sf(blue, mapping = aes(fill = name, colours = "blue")) +
  labs(title = "Glasgow blue places")

```


# Day 6 Red


```{r}
#  install.packages(c("devtools", "tidyverse"))
library(osmdata) 
library(sf)
library(tidyverse)

colours = "red"

colour <-  getbb(place_name = "glasgow uk") %>%
  opq() %>%
  add_osm_feature(key = "name", value = paste(colours), value_exact = FALSE, match_case = FALSE) %>% 
  osmdata_sf()

colour <- colour$osm_points
colour <- colour[grep(colours, colour$name, ignore.case = T), ]

colour$name <- as.factor(colour$name)
ggplot() +
  geom_sf(colour, mapping = aes(fill = name, colours = paste(colours))) +
  labs(title = "Glasgow red places") + theme(legend.position = "bottom")

```

# Day 7 Green


```{r}
#  install.packages(c("devtools", "tidyverse"))
library(osmdata) 
library(sf)
library(tidyverse)

colours = "green"

colour <-  getbb(place_name = "glasgow uk") %>%
  opq() %>%
  add_osm_feature(key = "name", value = paste(colours), value_exact = FALSE, match_case = FALSE) %>% 
  osmdata_sf()

colour <- colour$osm_points
colour <- colour[grep(colours, colour$name, ignore.case = T), ]

colour$name <- as.factor(colour$name)
ggplot() +
  geom_sf(colour, mapping = aes(fill = name, colours = paste(colours))) +
  labs(title = "Glasgow green places") + theme(legend.position = "bottom")

```

# Day 8 Yellow


```{r}
#  install.packages(c("devtools", "tidyverse"))
library(osmdata) 
library(sf)
library(tidyverse)

colours = "yellow"

colour <-  getbb(place_name = "glasgow uk") %>%
  opq() %>%
  add_osm_feature(key = "name", value = paste(colours), value_exact = FALSE, match_case = FALSE) %>% 
  osmdata_sf()

colour <- colour$osm_points
colour <- colour[grep(colours, colour$name, ignore.case = T), ]

colour$name <- as.factor(colour$name)
ggplot() +
  geom_sf(colour, mapping = aes(fill = name, colours = paste(colours))) +
  labs(title = "Glasgow yellow places") + theme(legend.position = "bottom")

```


# Day 9 Monochrome


```{r}
#  install.packages(c("devtools", "tidyverse", "osmdata"))
# devtools::install_github("fozy81/scotmaps")
library(scotmaps)
library(osmdata) 
library(sf)
library(tidyverse)


mono <-  getbb(place_name = "carluke scotland") %>%
  opq() %>%
  osmdata_sf()

 cc <- community_councils()
 carluke <- cc[cc$cc_name == "Carluke", ]

mono <- mono$osm_points
mono <- st_transform(mono, crs = 27700)

test <- st_intersection(mono, carluke)

plot(st_geometry(test), col = "black", cex = 0.01)


```


# Day 10 Grid

```{r}
#  install.packages(c("devtools", "tidyverse", "osmdata"))
# devtools::install_github("fozy81/scotmaps")
library(scotmaps)
library(osmdata) 
library(sf)
library(tidyverse)


la <- local_authorities()

cust_label <- function(x) {
  x <- abs(as.numeric(x))
  la$local_authority <- as.character(la$local_authority)
  la <- la %>% arrange(desc(hectares))
  paste0(la$local_authority[la$hectares %in% x])
}

ggplot(la) +
  geom_sf(fill = "lightgreen") +
  theme_void() +
  facet_wrap(vars(desc(hectares)),
             labeller = as_labeller(cust_label),
             strip.position = "bottom"
              ) +
  theme(
    strip.background = element_blank()
  ) +
  labs(title = "Scottish Local Authorities")



```


# Day 11: 3D

```{r}
# devtools::install_github("tylermorganwall/rayshader")
# devtools::install_github("datasciencescotland/opendatascot")
library(rayshader)
library(scotmaps)
library(opendatascot)
library(tidyverse)

# Immunisation data
la <- local_authorities()

immune <- ods_dataset("6-in-1-immunisation", geography = "la")
immune <- immune %>% filter(vaccinationUptake == "vaccinated-by-24-months", 
                     measureType == "ratio",
                     refPeriod == 2018)
immune <- immune %>% mutate(value = as.numeric(value))
# Join to local authority spatial layer
la_immune <- inner_join(la, immune, by = c("code" = "refArea"))

p2 <- la_immune %>% 
 ggplot() +
  geom_sf(aes(fill = value)) +
  scale_fill_viridis_c("Ratio") +
  theme_bw() +
  labs(title="6-in-1-immunisation by 24 months (2018)") +
  theme(plot.title = element_text(hjust = 1))



plot_gg(p2, width = 6, height=2.7)
render_camera(zoom = 0.4, theta = -45, phi = 30)
Sys.sleep(0.2)
render_snapshot()
rgl::rgl.close()


```

# Day 12: Not GIS software

#########x
#########x
#######x##
####xxx###
#x##xxx###
xx#xxxxxxx
x##xxxxxx#
###xxxxx##
#xxxxxx###
#x#xxxxxx#
#X#xxxx###
##xxxxxxx#


# Day 13: Raster

```{r, results="asis"}
library(osmdata)
library(leaflet)
library(tidyverse)

monuments <- getbb("carluke") %>% 
  opq() %>% 
  add_osm_feature(key = "historic", value = "memorial") %>% 
  osmdata_sf()

monuments$osm_points %>%
  head(5) %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = paste(monuments$osm_points$name))



```

# 14 Day: climate change

```{r}
# devtools::install_github("datasciencescotland/opendatascot")
library(opendatascot)
library(scotmaps)
library(tidyverse)
library(sf)

risk <- ods_dataset("dwellings-in-flood-risk-areas", geography = "la")

risk <- select(risk , floodRiskArea, value, refArea)
risk$value <- as.numeric(risk$value)
risk <- pivot_wider(risk, names_from =  floodRiskArea, values_from = value)

la <- local_authorities()

la <- inner_join(la, risk, by = c("code" = "refArea"))

plot(la[c("fluvial", "coastal", "coastal-or-fluvial")], main ="Flood Risk")

```


# Day 15: Connections

```{r}
#  install.packages(c("tidyverse", "osmdata"))
library(osmdata)
library(tidyverse)

ferry <- getbb("islay scotland") %>% 
  opq() %>% 
  
  add_osm_feature(key = "route", value = "ferry") %>% 
  osmdata_sf()


 ferry$osm_lines %>%
  leaflet() %>%
  addTiles() %>%
  addPolylines(popup = paste(ferry$osm_lines$name)) 

```

# Day 16: Islands

```{r}
#  install.packages(c("tidyverse", "osmdata"))
library(osmdata)
library(tidyverse)


islay <- getbb("islay uk") %>% 
  opq() %>% 
  add_osm_feature(key = "distillery", value = "whisky") %>% 
  osmdata_sf()


 islay$osm_polygons %>%
  leaflet() %>%
  addTiles() %>%
  addPolylines(popup = paste(islay$osm_polygons$name)) 

```

# Day 17: Historic

Historic Castles

```{r}
# install.packages(c("devtools", "tidyverse", "osmdata"))
library(osmdata)
library(tidyverse)

battles <- getbb("northumberland uk") %>% 
  opq() %>% 
  add_osm_feature(key = "historic", value = "castle") %>% 
  osmdata_sf()

 battles$osm_polygons %>%
  leaflet() %>%
  addTiles() %>%
  addPolylines(popup = paste(battles$osm_polygons$name)) 

```

# Day 18: Landuse 

```{r}
# install.packages(c("tidyverse", "osmdata"))
library(osmdata)
library(tidyverse)

kelso <- getbb("kelso scotland") %>% 
  opq() %>% 
  add_osm_feature(key = "natural") %>% 
  osmdata_sf()

kelso <- kelso$osm_polygons

ggplot() +
  geom_sf(kelso, mapping = aes(colour = natural, fill = natural)) +
  labs(title = "Natural Landuse Kelso") +
  theme_void() 


```

# Day 19: NULL

```{r}
# install.packages(c("tidyverse", "osmdata"))
library(osmdata)
library(sf)

null <- 
  opq(bbox = c(0.00,0.00,0.00,0.00)) %>% 
  osmdata_sf() 
  

plot(st_geometry(null$osm_points), main = "NULL")

```


# Day 20 Population

```{r}
# devtools::install_github("datasciencescotland/opendatascot")
# install.packages(c("tidyverse", "osmdata"))
library(scotmaps)
library(opendatascot)
library(tidyverse)

datasets <- ods_all_datasets()
ods_structure("population-projections-2018-based")

pop <- ods_dataset("population-projections-2018-based",
                   geography = "la",
                   refPeriod = "2043")

pop20 <- ods_dataset("population-projections-2018-based",
                   geography = "la",
                   refPeriod = "2020")

pop$value <- as.numeric(pop$value)
pop20$value <- as.numeric(pop20$value)
 pop$diff  <- 1 /  pop$value *  (pop$value - pop20$value)

test <- filter(pop, sex == "all", populationProjectionVariant == "principal-projection") %>% 
  mutate(refArea = as.factor(refArea), value = as.numeric(diff)) %>% 
  select(refArea, diff) %>% 
  group_by(refArea) %>% 
  summarise(total = sum(diff))

la <- local_authorities()

la <- inner_join(la, test, by = c("code" = "refArea"))

ggplot() +
  geom_sf(la, mapping = aes(fill = total)) +
  labs(title = "2020-43 Predicted Population % Change")

```

# Day 21: Water


River Clyde

```{r, results="asis"}
# install.packages(c("tidyverse", "osmdata", "leaflet"))
library(osmdata)
library(tidyverse)
library(leaflet)

river <- 
  opq(bbox = "scotland") %>% 
  add_osm_feature(key = "waterway", value = "river") %>% 
  add_osm_feature(key = "name", value = "River Clyde" ) %>% 
  osmdata_sf() 
  
 river$osm_lines %>% 
 leaflet() %>% 
   addTiles() %>% 
   addPolylines()
 

```


# Day 22: Movement

Marinas around island of Ireland
```{r}
# install.packages(c("tidyverse", "osmdata"))
library(osmdata)
library(tidyverse)

marina <- 
  opq(bbox = "Ireland") %>% 
  add_osm_feature(key = "leisure", value = "marina") %>% 
  osmdata_sf() 
  
plot(st_geometry(marina$osm_polygons))
plot(st_geometry(marina$osm_multipolygons), add = TRUE)
plot(st_geometry(marina$osm_points), add = TRUE, col = "darkgreen")

```


# Day 23 Boundaries

```{r}
# install.packages(c("tidyverse", "osmdata"))
library(osmdata)
library(tidyverse)

border <-
opq(bbox = "Baarle-Hertog") %>% 
  add_osm_feature(key = "boundary", value = "administrative") %>% 
   add_osm_feature(key = "border_type", value = "nation") %>% 
  osmdata_sf()

plot(st_geometry(border$osm_polygons))
plot(st_geometry(border$osm_lines), add=TRUE)
plot(st_geometry(border$osm_multilines), add=TRUE)

```


# Day 24: Elevation




